---
title: "Assignment II: Group [A]"
author: "HE Bingyi"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: HE Bingyi**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
xxxxxxxxxxxxxx
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

BIOTIME_raw_data_232.csv
:::

```{r}
bio <- read.csv("data/BIOTIME_raw_data_232.csv")
head(bio)
str(bio)
```

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

#### 1.1 Dataset Summary

The dataset used in this study comes from the **BioTIME database (Study ID 232)**. Rather than representing general marine records, this dataset specifically focuses on **biological surveys conducted in the waters surrounding the sub-Antarctic Kerguelen Islands**. It contains ecological observations of marine organisms and is useful for studying biodiversity patterns in a high-latitude marine ecosystem.

#### 1.2 Temporal Scope

The dataset mainly covers the period from **1990 to 1992**. In addition to yearly information, the data also include observations recorded by **month (`MONTH`)** and **day (`DAY`)**, which makes it possible to examine **short-term and seasonal fluctuations** in abundance and biomass during the sampling period.

#### 1.3 Taxonomic Coverage

The dataset includes multiple marine taxa, with several representative species appearing repeatedly in the observations. One of the most frequently recorded species is ***Mancopsetta maculata***, a fish species that appears to play an important role in this dataset. Other taxa, such as members of the **Zoarcidae** family, are also present.

These species are important components of the **Southern Ocean food web**, and changes in their **abundance (`ABUNDANCE`)** and **biomass (`BIOMAS`)** may reflect shifts in the ecological condition and stability of the local marine ecosystem.

#### 1.4 Spatial Distribution

The sampling locations are concentrated in the waters between approximately **51°S and 58°S latitude**, and **72°E and 77°E longitude**. This specific geographic region is strongly influenced by **Southern Ocean circulation and environmental conditions**, making it a valuable area for studying **marine biodiversity and species distribution in sub-Antarctic ecosystems**.

#### 1.5 Variables Included in the Dataset

This dataset contains **6,349 observations and 9 variables**. The main variables include:

-   **`ABUNDANCE`**: the number of individuals recorded in a sampling event\
-   **`BIOMAS`**: the biomass associated with the recorded observation\
-   **`valid_name`**: the scientific name of the species\
-   **`LATITUDE`** and **`LONGITUDE`**: the geographic coordinates of the sampling site\
-   **`DAY`**, **`MONTH`**, and **`YEAR`**: the date of sampling\
-   **`SAMPLE_DESC`**: a sample identifier that appears to combine spatial and temporal information

Overall, this dataset is suitable for visualization because it contains **species-level, temporal, and spatial information**, allowing us to explore how marine biodiversity patterns vary across **time, species, and location**.
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

I chose an **interactive scatter plot** created using the `plotly` package to visualize the relationship between **Abundance** and **Biomass** for the top 3 most abundant species in the dataset. In addition to one combined plot, I also created three separate species-specific plots so that the overall pattern and the individual species patterns can both be examined clearly.

I selected this plot type because a scatter plot is one of the most suitable ways to display the relationship between **two continuous variables**. In this dataset, **Abundance** represents how many individuals were recorded, while **Biomass** represents the total biological mass. Plotting these two variables together helps reveal whether higher abundance is generally associated with higher biomass and whether different species follow similar or different ecological patterns.

The main **strength** of this visualization is that it is **interactive**, which provides clear advantages over a static plot. In the combined plot, users can hover over each point to inspect detailed information such as species name, abundance, biomass, and year. They can also compare the three species more easily through color coding. The separate plots further improve readability because they reduce overlap and make within-species patterns easier to observe.

Another strength is that the scatter plot is visually intuitive. It clearly shows clusters, trends, and extreme observations (outliers), which helps readers understand the ecological structure of the dataset more effectively.

However, this plot also has some **weaknesses**. First, when many points are concentrated in a small area, some overlap still occurs even after adding slight jitter. Second, the scatter plot mainly provides a **visual** interpretation and does not directly give a statistical measurement such as a correlation coefficient or regression result. Finally, because biomass values vary greatly, the plot may be influenced by several high-value observations.

Overall, this interactive scatter plot is an effective and meaningful way to explore species-level ecological patterns in the dataset.
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

The purpose of this visualization is to explore and communicate the relationship between **species abundance and biomass** in the selected marine biodiversity dataset.

More specifically, this visualization is designed to show whether species with higher abundance also tend to have higher biomass, and whether this relationship differs among the top 3 most abundant species. By using one combined plot together with separate species-specific plots, the figure allows readers to first understand the **overall ecological pattern** and then examine each species in more detail.

This visualization also helps identify **clusters, variation, and possible outliers** in the data. These patterns are important because they may reflect species-specific ecological characteristics, differences in body size, or different abundance structures within the marine ecosystem.

The interactive features make the visualization more useful than a static figure because readers can inspect individual observations and compare species more flexibly. Therefore, the figure is intended not only to present the data correctly, but also to help readers better understand the ecological meaning behind the observed patterns.
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
# YOUR CODE HERE
# xxxxx


#| echo: true

library(dplyr)
library(plotly)

# -----------------------------
# Step 1: Clean data
# -----------------------------
bio_clean <- bio %>%
  filter(!is.na(ABUNDANCE),
         !is.na(BIOMAS),
         !is.na(valid_name),
         ABUNDANCE > 0,
         BIOMAS > 0)

# -----------------------------
# Step 2: Top 3 species
# -----------------------------
top3 <- bio_clean %>%
  group_by(valid_name) %>%
  summarise(total_abundance = sum(ABUNDANCE, na.rm = TRUE)) %>%
  arrange(desc(total_abundance)) %>%
  slice_head(n = 3)

bio_top3 <- bio_clean %>%
  filter(valid_name %in% top3$valid_name) %>%
  mutate(
    ABUNDANCE_jitter = jitter(ABUNDANCE, amount = 2),
    BIOMAS_jitter = jitter(BIOMAS, amount = 300)
  )

species_names <- unique(bio_top3$valid_name)

# -----------------------------
# Plot 1: Combined plot
# -----------------------------
plot_ly(
  bio_top3,
  x = ~ABUNDANCE_jitter,
  y = ~BIOMAS_jitter,
  color = ~valid_name,
  colors = c(
    "Champsocephalus gunnari" = "#66c2a5",
    "Channichthys rhinoceratus" = "#fc8d62",
    "Pleuragramma antarcticum" = "#8da0cb"
  ),
  type = "scatter",
  mode = "markers",
  marker = list(size = 8, opacity = 0.75),
  text = ~paste(
    "Species:", valid_name,
    "<br>Abundance:", ABUNDANCE,
    "<br>Biomass:", BIOMAS,
    "<br>Year:", YEAR
  ),
  hoverinfo = "text"
) %>%
  layout(
    title = "Abundance and Biomass of the Top 3 Species",
    xaxis = list(title = "Abundance"),
    yaxis = list(title = "Biomass"),
    legend = list(title = list(text = "Species"))
  )

# -----------------------------
# Plot 2: Champsocephalus gunnari
# -----------------------------
plot_ly(
  data = filter(bio_top3, valid_name == "Champsocephalus gunnari"),
  x = ~ABUNDANCE_jitter,
  y = ~BIOMAS_jitter,
  type = "scatter",
  mode = "markers",
  marker = list(
    color = "#66c2a5",
    size = 8,
    opacity = 0.8
  ),
  text = ~paste(
    "Species:", valid_name,
    "<br>Abundance:", ABUNDANCE,
    "<br>Biomass:", BIOMAS,
    "<br>Year:", YEAR
  ),
  hoverinfo = "text"
) %>%
  layout(
    title = "Champsocephalus gunnari",
    xaxis = list(title = "Abundance"),
    yaxis = list(title = "Biomass")
  )

# -----------------------------
# Plot 3: Channichthys rhinoceratus
# -----------------------------
plot_ly(
  data = filter(bio_top3, valid_name == "Channichthys rhinoceratus"),
  x = ~ABUNDANCE_jitter,
  y = ~BIOMAS_jitter,
  type = "scatter",
  mode = "markers",
  marker = list(
    color = "#fc8d62",
    size = 8,
    opacity = 0.8
  ),
  text = ~paste(
    "Species:", valid_name,
    "<br>Abundance:", ABUNDANCE,
    "<br>Biomass:", BIOMAS,
    "<br>Year:", YEAR
  ),
  hoverinfo = "text"
) %>%
  layout(
    title = "Channichthys rhinoceratus",
    xaxis = list(title = "Abundance"),
    yaxis = list(title = "Biomass")
  )

# -----------------------------
# Plot 4: Pleuragramma antarcticum
# -----------------------------
plot_ly(
  data = filter(bio_top3, valid_name == "Pleuragramma antarcticum"),
  x = ~ABUNDANCE_jitter,
  y = ~BIOMAS_jitter,
  type = "scatter",
  mode = "markers",
  marker = list(
    color = "#8da0cb",
    size = 8,
    opacity = 0.8
  ),
  text = ~paste(
    "Species:", valid_name,
    "<br>Abundance:", ABUNDANCE,
    "<br>Biomass:", BIOMAS,
    "<br>Year:", YEAR
  ),
  hoverinfo = "text"
) %>%
  layout(
    title = "Pleuragramma antarcticum",
    xaxis = list(title = "Abundance"),
    yaxis = list(title = "Biomass")
  )

```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

```{r}
# YOUR CODE HERE
# xxxxx



#| echo: true
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)
library(gganimate)
library(forcats)
library(scales)

bio <- read.csv("data/BIOTIME_raw_data_232.csv")

bio_clean <- bio %>%
  filter(!is.na(valid_name),
         !is.na(ABUNDANCE),
         !is.na(YEAR),
         ABUNDANCE > 0)

top8 <- bio_clean %>%
  group_by(valid_name) %>%
  summarise(total_abundance = sum(ABUNDANCE, na.rm = TRUE)) %>%
  arrange(desc(total_abundance)) %>%
  slice_head(n = 8)

anim_data <- bio_clean %>%
  filter(valid_name %in% top8$valid_name) %>%
  group_by(YEAR, valid_name) %>%
  summarise(yearly_abundance = sum(ABUNDANCE, na.rm = TRUE), .groups = "drop")

p_anim <- ggplot(
  anim_data,
  aes(
    x = yearly_abundance,
    y = fct_reorder(valid_name, yearly_abundance),
    fill = valid_name
  )
) +
  geom_col(show.legend = FALSE, width = 0.8) +
  coord_flip() +
  labs(
    title = "Top 8 Fish Species by Abundance in Year: {closest_state}",
    x = "Total Abundance",
    y = "Species"
  ) +
  scale_x_continuous(labels = comma) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 11)
  )

p_anim +
  transition_states(
    YEAR,
    transition_length = 2,
    state_length = 1
  ) +
  ease_aes("cubic-in-out")
```

```{r}

#| echo: true
#| warning: false
#| message: false

library(dplyr)
library(plotly)

# Step 1: read data
bio <- read.csv("data/BIOTIME_raw_data_232.csv")

# Step 2: clean data
bio_clean <- bio %>%
  filter(!is.na(valid_name),
         !is.na(ABUNDANCE),
         !is.na(BIOMAS),
         !is.na(YEAR),
         ABUNDANCE > 0,
         BIOMAS > 0)

# Step 3: select top 3 species by total abundance
top3 <- bio_clean %>%
  group_by(valid_name) %>%
  summarise(total_abundance = sum(ABUNDANCE, na.rm = TRUE)) %>%
  arrange(desc(total_abundance)) %>%
  slice_head(n = 3)

anim_data <- bio_clean %>%
  filter(valid_name %in% top3$valid_name)

# Step 4: animated scatter plot
plot_ly(
  data = anim_data,
  x = ~ABUNDANCE,
  y = ~BIOMAS,
  color = ~valid_name,
  colors = c(
    "Champsocephalus gunnari" = "#66c2a5",
    "Channichthys rhinoceratus" = "#fc8d62",
    "Pleuragramma antarcticum" = "#8da0cb"
  ),
  frame = ~YEAR,
  type = "scatter",
  mode = "markers",
  marker = list(size = 10, opacity = 0.7),
  text = ~paste(
    "Species:", valid_name,
    "<br>Abundance:", ABUNDANCE,
    "<br>Biomass:", BIOMAS,
    "<br>Year:", YEAR
  ),
  hoverinfo = "text"
) %>%
  layout(
    title = "Animated Abundance-Biomass Patterns of the Top 3 Fish Species",
    xaxis = list(title = "Abundance"),
    yaxis = list(title = "Biomass"),
    legend = list(title = list(text = "Species"))
  ) %>%
  animation_opts(
    frame = 1000,
    transition = 500,
    redraw = FALSE
  ) %>%
  animation_slider(
    currentvalue = list(prefix = "Year: ")
  ) %>%
  animation_button(
    x = 0.1,
    xanchor = "right",
    y = 0,
    yanchor = "top"
  )
```

```{r}
# --- 1. 加载包 ---
library(tidyverse)
library(plotly)
library(crosstalk)

# --- 2. 准备 Shared Data (实现颗粒并行联动的关键) ---
# 选取前 10 个物种，不汇总，保留所有原始采样颗粒
top_species <- bio %>%
  count(valid_name, sort = TRUE) %>%
  head(10) %>%
  pull(valid_name)

plot_data <- bio %>%
  filter(valid_name %in% top_species) %>%
  filter(ABUNDANCE > 0 & BIOMAS > 0) %>%
  drop_na(ABUNDANCE, BIOMAS, YEAR)

# 使用 crosstalk 创建共享数据集
shared_bio <- highlight_key(plot_data, ~valid_name)

# --- 3. 创建两个平行的基础图 ---
# 图 A: 丰度随年份分布 (颗粒点)
p1 <- ggplot(shared_bio, aes(x = YEAR, y = ABUNDANCE, color = valid_name)) +
  geom_jitter(alpha = 0.4, size = 1) + 
  scale_y_log10() +
  theme_minimal() +
  labs(y = "Abundance (Log Scale)", x = "") +
  theme(legend.position = "none")

# 图 B: 生物量随年份分布 (颗粒点)
p2 <- ggplot(shared_bio, aes(x = YEAR, y = BIOMAS, color = valid_name)) +
  geom_jitter(alpha = 0.4, size = 1) + 
  scale_y_log10() +
  theme_minimal() +
  labs(y = "Biomass (Log Scale)", x = "Year") +
  theme(legend.position = "none")

# --- 4. 合并为并行子图 ---
# 两个变量在每一年（X轴）同时呈现，完全满足你的并行要求
subplot(ggplotly(p1), ggplotly(p2), nrows = 2, shareX = TRUE, titleY = TRUE) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick", color = "red") %>%
  layout(title = "Parallel Trends: Abundance & Biomass (1968-1999)")
```

```{r}
library(tidyverse)
library(plotly)

# 1. 读入数据
bio <- read.csv("data/BIOTIME_raw_data_232.csv")

# 2. 找出总 abundance 最高的前 10 个物种
top10_species <- bio %>%
  group_by(valid_name) %>%
  summarise(total_abundance = sum(ABUNDANCE, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total_abundance)) %>%
  slice_head(n = 10) %>%
  pull(valid_name)

# 3. 保留前10物种的原始记录，不做 yearly summary
plot_data <- bio %>%
  filter(valid_name %in% top10_species) %>%
  filter(!is.na(ABUNDANCE), !is.na(BIOMAS), !is.na(YEAR)) %>%
  filter(ABUNDANCE > 0, BIOMAS > 0)

# 4. 画图：保持老师第一段的结构
base2 <- plot_ly(
  data = plot_data,
  x = ~ABUNDANCE,
  y = ~BIOMAS,
  color = ~valid_name,
  colors = "Set2"
) %>%
  add_markers(
    marker = list(size = 5, opacity = 0.18),
    showlegend = TRUE,
    hoverinfo = "skip"
  ) %>%
  add_markers(
    frame = ~YEAR,
    marker = list(size = 8, opacity = 1),
    hovertemplate = paste(
      "Species: %{customdata}<br>",
      "Abundance: %{x}<br>",
      "Biomass: %{y}<br>",
      "Year: %{frame}<extra></extra>"
    ),
    customdata = ~valid_name
  ) %>%
  layout(
    xaxis = list(title = "Abundance", type = "log"),
    yaxis = list(title = "Biomass", type = "log"),
    legend = list(title = list(text = "Species"))
  ) %>%
  animation_opts(1000, redraw = FALSE) %>%
  animation_slider(currentvalue = list(prefix = "YEAR: "))

base2
```
