---
title: "Assignment II: Group A"
author: "YANG,Taoming"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: YANG, Taoming**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
 During the assignment, I used Deepseek to assist me in completing it. Deepseek specifically helped me in two ways: 1. by polishing comments; 2. by improving code functionality. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

ggplot2::msleep
:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

This data records details of mammalian sleep and species habits. The dataset contains 83 mammal species and 11 variables. The following is a description of the variables.\
\
name: Species name\
genus: Genus name\
vore: Diet\
order: Order name\
conservation: Conservation status\
sleep_total: Total sleep time in hours\
sleep_rem: Rapid eye movement sleep time in hours\
sleep_cycle: Sleep cycle in minutes\
awake: Awake time in hours\
brainwt: Brain weight in grams\
bodywt: Body weight in kilograms\
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

plot_1: Sleep Total Time Distribution of Mammals: \
Pros and Cons: I chose an interactive histogram to display the distribution of sleep duration in mammals. The advantage is that the interactive histogram provides an overview of data distribution and quickly identifies outliers; hovering the mouse shows the count for each bin, and zooming is easy.\

plot_2: Mammalian Diet vs Sleep Total Time: \
Pros and Cons: I chose a boxplot supplemented with jittered scatter points to observe data distribution and density while avoiding hidden data points due to overplotting. The disadvantage is that in areas of high data density, points may still be obscured, and in some cases, the distance between quartiles and extremes is too small, making details unclear.\

plot_3: Body Size vs Sleep Duration Relationship: \
Pros and Cons: I used an interactive scatter plot to show the relationship between body weight and sleep duration, with a logarithmic transformation applied to the x-axis.The interactive scatter plot handles discrete data points intuitively and allows direct observation of the relationship between two variables. Hovering the mouse reveals data for each point. The disadvantage is that the x-axis uses a log transformation, which may make the actual trends appear more extreme than shown, posing a reading challenge for some viewers.\
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

plot_1:\
Purpose: To quickly understand the distribution of total sleep time in the initial stage of research and confirm whether there are outliers that need to be removed.\

plot_2:\
Purpose: To analyze whether there are significant differences in sleep duration among mammals with different diets.\

plot_3:\
Purpose: I used an interactive scatter plot to show the relationship between body weight and sleep duration, with a logarithmic transformation applied to the x-axis.
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
# Load necessary packages
library(tidyverse)
library(plotly)
library(ggplot2)
library(dplyr)
library(scales)
library(gganimate)

# Load data
data("msleep")
dim(msleep)
head(msleep)
str(msleep)

# Check the percentage of NAs for each variable
apply(msleep,2,FUN = function(x) sum(is.na(x))/dim(msleep)[1])



# 1. Plot the overall distribution of sleep total time in mammals

g1 <- ggplot(msleep, aes(x = sleep_total)) +
  geom_histogram(binwidth = 2, fill = "steelblue", color = "black") +
  theme_minimal() +
  labs(title = "Sleep Total Time Distribution of Mammals",
       x = "Sleep Total Time (h)",
       y = "Count") +
  scale_x_continuous(breaks = seq(1, 21, by = 2)) # Add appropriate x-axis scale

# Convert to interactive plotly object
ggplotly(g1)


# 2. Boxplot + jittered scatter plot of sleep duration by diet type in mammals

# Data processing
## Since there are a few missing values in the vore variable, remove samples with missing values before plotting
msleep_1 <- msleep[!is.na(msleep$vore),]

g2 <- ggplot(msleep_1, aes(x = vore, y = sleep_total, color = vore)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height = 0, width = 0.2, size = 2, alpha = 0.8) +
  theme_minimal() +
  labs(title = "Mammalian Diet vs Sleep Total Time",
       x = "Mammalian Diet Type",
       y = "Sleep Total Time (h)")

ggplotly(g2)


# 3. Plot body size vs sleep duration relationship     

g3 <- ggplot(msleep, aes(x = bodywt, y = sleep_total)) +
  geom_point() +
  scale_x_log10(labels = comma) + # Log transformation for x-axis
  theme_minimal() +
  labs(title = "Mammalian Body Weight vs Sleep Total Time",
       x = "log(Mammalian Body Weight)",
       y = "Sleep Total Time (h)")

ggplotly(g3)

```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

plot_4: Distribution of body weight and total sleep duration among mammals with different diets:\
Pros and cons: I chose ggplot with gganimate's transition animation. The advantage of transition animation is that it clearly distinguishes different groups, displaying only a single group at a time. Compared to static or interactive plots, it is more intuitive and engaging, while enhancing comparison between groups. The disadvantages are longer loading times and higher resource consumption; groups must be viewed sequentially, making it impossible to see all information at once; exact data values cannot be extracted; and it cannot be displayed in static documents.
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

The purpose is to further explore whether new patterns or characteristics emerge in the body weight vs sleep duration scatter plot when grouped by diet category.
:::

```{r}
#| echo: true

# 4. Distribution of body weight and total sleep duration among mammals with different diets

# The plot part
a1 <- ggplot(msleep_1, aes(x = bodywt, y = sleep_total)) + 
  scale_x_log10(labels = comma) +
  geom_point(aes(colour = vore), size = 4) + 
  theme_minimal() +
  labs(title = "Body Weight vs Total Sleep Time by Diet in Mammals",
       x = "log(Mammalian Body Weight)",
       y = "Sleep Total Time (h)") +
  transition_states(vore,
                    transition_length = 2,
                    state_length = 1)

# The animation part
a1 + 
  enter_fade() + enter_drift(x_mod = -1) + 
  exit_shrink() + exit_drift(x_mod = 4)
```
